---
title: Armed Strength- Analyzing National Capabilities and Conflict Frequency in Asia
subtitle: Report
execute:
  echo: false
---

```{r}
#| label: load-pkgs
#| message: false
#| echo: false
library(tidyverse)
library(readxl)
library(countrycode) 
library(tidymodels)
library(openintro)
library(ggcorrplot)
library(broom)
library(kableExtra)

#main dataset abt resources

country_resource <- read_csv("data/NMC.csv")

#main dataset abt conflict

country_conflict <- read_csv("data/Conflict.csv")

```

## Introduction and Data

<!--- cleaning -->

```{r}
#| label: taking-out-unnecessary-variables
#| echo: false

country_resource <- country_resource |>
  select(-ccode, -version)

country_conflict <- country_conflict |>
  select(-ccode, -version)
```

```{r}
#| label: join-datasets
#| echo: false
country_join <- country_conflict |>
  full_join(country_resource, by = join_by(
      stabb == stateabb, 
      styear == year
    )
  )

number_of_disputes_df <- country_join |>
  group_by(stabb, styear) |>
  summarize(count = n(), .groups = "drop")


```

```{r}
#| label: filter-dataset
#| echo: false
country_join_edit <- country_join |>
  left_join(number_of_disputes_df, by = c("stabb" = "stabb", "styear" = "styear"))

country_join_edit <- country_join |>
  left_join(number_of_disputes_df, by = c("stabb" = "stabb", "styear" = "styear"))

country_join_shorter <- country_join_edit |>
  select(stabb, styear, count, milex, milper, irst, cinc) |>
  distinct() |>
  filter(styear>1950)
```

```{r}
#| label: continents 
#| message: false
#| warning: false
#| echo: false

# Manually addded countried from the COW dataset, COW codes do not match universal country codes so there were many NAs using the country code package to use the continents 

# cow code to country name --> to actual ISO code --> continents 

#stabb_values <- country_join_edit[["stabb"]] isolating COW code
#unique_stabb_values <- unique(stabb_values) removing duplicates; 219 countries unique 



CountryName <- data.frame(
  COWn = c("ZIM", "ZAN", "ZAM", "YUG", "YPR", "YEM", "YAR", "WSM", "WRT", "VEN",
           "VAN", "UZB", "USA", "URU", "UKR", "UKG", "UGA", "UAE", "TUV", "TUS",
           "TUR", "TUN", "TRI", "TON", "TOG", "TKM", "THI", "TAZ", "TAW", "TAJ", 
           "SYR", "SWZ", "SWD", "SWA", "SVG", "SUR", "SUD", "STP", "SSD", "SRI", 
           "SPN", "SOM", "SOL", "SNM", "SLV", "SLU", "SLO", "SKN", "SIN", "SIE", 
           "SIC", "SEY", "SEN", "SAX", "SAU", "SAL", "SAF", "RWA", "RVN", "RUS", 
           "ROM", "ROK", "QAT", "PRK", "POR", "POL", "PNG", "PMA", "PHI", "PER", 
           "PAR", "PAP", "PAN", "PAL", "PAK", "OMA", "NTH", "NOR", "NIR", "NIG", 
           "NIC", "NEW", "NEP", "NAU", "NAM", "MZM", "MYA", "MSI", "MOR", "MON", 
           "MOD", "MNG", "MNC", "MLT", "MLI", "MLD", "MEX", "MEC", "MAW", "MAS", 
           "MAL", "MAG", "MAD", "MAC", "MAA", "LUX", "LIT", "LIE", "LIB", "LES", 
           "LEB", "LBR", "LAT", "LAO", "KZK", "KYR", "KUW", "KOS", "KOR", "KIR", 
           "KEN", "JPN", "JOR", "JAM", "ITA", "ISR", "IRQ", "IRN", "IRE", "INS", 
           "IND", "ICE", "HUN", "HSG", "HSE", "HON", "HKG", "HAN", "HAI", "GUY", 
           "GUI", "GUA", "GRN", "GRG", "GRC", "GNB", "GMY", "GHA", "GFR", "GDR",
           "GAM", "GAB", "FSM", "FRN", "FIN", "FIJ", "EU", "ETM", "ETH", "EST", 
           "ERI", "EQG", "EGY", "ECU", "DRV", "DRC", "DOM", "DMA", "DJI", "DEN", 
           "CZR", "CZE", "CYP", "CUB", "CRO", "COS", "CON", "COM", "COL", 
           "CHN", "CHL", "CHA", "CEN", "CDI", "CAP", "CAO", "CAN", "CAM", 
           "BUL", "BUI", "BRU", "BRA", "BOT", "BOS", "BOL", "BNG", "BLZ", "BLR",
           "BHU", "BHM", "BFO", "BEN", "BEL", "BAV", "BAR", "BAH", "BAD", "AZE", 
           "AUS", "AUL", "AUH", "ARM", "ARG", "ANG", "AND", 
           "ALG", "ALB", "AFG"),
  
  
  Country = c("Zimbabwe", "Zanzibar", "Zambia", "Yugoslavia/Serbia",
              "Yemen People's Republic", "Yemen", "Yemen Arab Republic", 
              "Samoa", "Wuerttemburg", "Venezuela", "Vanuatu", "Uzbekistan", 
              "United States of America", "Uruguay", "Ukraine", 
              "United Kingdom", "Uganda", "United Arab Emirates", "Tuvalu", 
              "Tuscany", "Turkey", "Tunisia", "Trinidad and Tobago", "Tonga", 
              "Togo", "Turkmenistan", "Thailand", "Tanzania", "Taiwan",
              "Tajikistan", "Syria", "Switzerland", "Sweden", "Swaziland", 
              "St. Vincent and the Grenadines", "Suriname", "Sudan",
              "Sao Tome and Principe", "South Sudan", "Sri Lanka", "Spain", 
              "Somalia", "Solomon Islands", "San Marino", "Slovenia", 
              "St. Lucia", "Slovakia", "St. Kitts and Nevis", "Singapore", 
              "Sierra Leone", "Two Sicilies", "Seychelles", "Senegal", 
              "Saxony", "Saudi Arabia", "El Salvador", "South Africa",
              "Rwanda", "Republic of Vietnam", "Russia", "Romania", 
              "South Korea", "Qatar", "North Korea", "Portugal", "Poland", 
              "Papua New Guinea", "Parma", "Philippines", "Peru", "Paraguay",
              "Papal States", "Panama", "Palau", "Pakistan", "Oman", 
              "Netherlands", "Norway", "Niger", "Nigeria", "Nicaragua", 
              "New Zealand", "Nepal", "Nauru", "Namibia", "Mozambique",
              "Myanmar", "Marshall Islands", "Morocco", "Mongolia", "Modena", 
              "Montenegro", "Monaco", "Malta", "Mali", "Moldova", 
              "Mexico", "Mecklenburg Schwerin", "Malawi", "Mauritius", 
              "Malaysia", "Madagascar", "Maldives", "Macedonia", "Mauritania", 
              "Luxembourg", "Lithuania", "Liechtenstein", "Libya", "Lesotho",
              "Lebanon", "Liberia", "Latvia", "Laos", "Kazakhstan", 
              "Kyrgyzstan", "Kuwait", "Kosovo", "Korea", "Kiribati", 
              "Kenya", "Japan", "Jordan", "Jamaica", "Italy", "Israel", "Iraq",
              "Iran", "Ireland", "Indonesia", "India", "Iceland", "Hungary",
              "Hesse Grand Ducal", "Hesse Electoral", "Honduras", "Hong Kong", 
              "Hanover", "Haiti", "Guyana", "Guinea", "Guatemala", "Grenada", 
              "Georgia", "Greece", "Guinea-Bissau", "Germany", "Ghana", 
              "German Federal Republic", "German Democratic Republic", 
              "Gambia", "Gabon", "Federated States of Micronesia", "France", 
              "Finland", "Fiji", "European Community/Union", "East Timor",
              "Ethiopia", "Estonia", "Eritrea", "Equatorial Guinea", 
              "Egypt", "Ecuador", "Vietnam", "Democratic Republic of the Congo",
              "Dominican Republic", "Dominica", "Djibouti", "Denmark", 
              "Czech Republic", "Czechoslovakia", "Cyprus", "Cuba", 
              "Croatia", "Costa Rica", "Congo", "Comoros", "Colombia", "China",
              "Chile", "Chad", "Central African Republic", "Ivory Coast", 
              "Cape Verde", "Cameroon", "Canada", "Cambodia", "Bulgaria", 
              "Burundi", "Brunei", "Brazil", "Botswana", 
              "Bosnia and Herzegovina", "Bolivia", "Bangladesh", "Belize", 
              "Belarus", "Bhutan", "Bahamas", "Burkina Faso", "Benin", 
              "Belgium", "Bavaria", "Barbados", "Bahrain", "Baden",
              "Azerbaijan", "Austria", "Australia", "Austria-Hungary", 
              "Armenia", "Argentina", "Angola", "Andorra", "Algeria", 
              "Albania", "Afghanistan")
)

country_join_countrynames <- country_join_shorter |>
  left_join(CountryName, by = c("stabb" = "COWn")) 


# country column to get the ISO codes and continents
country_join_continents <- country_join_countrynames |>
  mutate(
    # join for country code conversion
    Country_code = countrycode(Country, "country.name", "iso3c"),
    # convert the ISO country code to continent
    Continent = countrycode(Country_code, "iso3c", "continent")
  ) |>
  filter(!is.na(Country)) |>
  filter(!is.na(Continent))

#fliter by asia 
asia <- country_join_continents |>
  filter(Continent == "Asia")

```

The primary motivation for this research is to explore the relationship between national material capabilities and a country’s likelihood of being involved in conflicts.
Understanding how changes in a nation’s military capabilities — such as military expenditures and natural resources like iron — affect its involvement in international disputes is important for assessing political stability.

The data used in our analysis comes from the [Correlates of War (COW) project](https://correlatesofwar.org/data-sets/), which consists of two datasets that apply to our research- [National Material Capabilities](https://correlatesofwar.org/data-sets/national-material-capabilities/) and [Militarized Interstate Disputes](https://correlatesofwar.org/data-sets/mids/).
The former dataset includes annually collected data on military power indicators for each nation, such as military expenditures, military personnel, etc.
The latter records instances where states threaten, display, or use force against one another every year from 1816 to 2014.

After joining these two datasets, we mutated our dataset to include the variables: “count” (summing up the total incident count for a given country in a given year), “milex_class” (classifying the military expenditure as being either above or below the median military expenditure), and “irst_class”(classifying the iron and steel production as being either above or below the median iron and steel Production).
In preparing the data, we addressed several challenges.
Variables irrelevant to our analysis were removed, and we were forced to correct the unique country codes used in the dataset as they didn’t align with standard naming conventions.
These codes were manually mapped to their respective countries (so as to not introduce a new dataset), and we classified nations, adding new variables, country and continent, to narrow the scope of analysis.

@tbl-dict provides an overview of relevant variables used in this analysis:

```{r}
#| label: tbl-dict
#| echo: false
#| tbl-cap: "Excerpt of Data Dictionary"


variable_info <- tibble::tribble(
  ~Variable,      ~Definition,                                                                 ~Type,
  "country",      "Nation of observation",                         "categorical",
  "continent",    "Geographic area of the country (e.g., Asia)",                 "categorical",
  "count",        "Total number of incidents of conflict",                    "numerical",
  "milex",        "Military expenditures in thousands in international currency",      "numerical",
  "irst",         "Iron and Steel production levels in thousands of tons",             "numerical"
)

variable_info |>
kable()
```

**Our research question is as follows:** How do changes in national material capabilities influence the frequency of a country's conflicts, focusing on the continent of Asia?

**We hypothesize** that countries with higher military expenditures and iron and steel production are more likely to participate in armed conflicts.

This research aligns with a broader theme of examining how data can explain political dynamics, provide insights into state behavior, and potentially help predict future conflicts.
In analyzing the intersection of politics and data, this study aims to tell a story about how power, in its different forms, shapes international relations.

In order to narrow our search for proper analyses, we decided to focus on the continent with the highest propensity for conflict.
@fig-high-prop-conflict identifies Asia as the continent with the highest propensity for conflict.

```{r}
#| label: fig-high-prop-conflict 
#| fig-cap: "Continent with the Highest Propensity for Conflict"
#| message: false
#| warning: false
#| echo: false

conflict_by_continent <- country_join_continents |>
  group_by(Continent) |>
  summarize(total_conflict = sum(count, na.rm = TRUE)) |>
  arrange(desc(total_conflict))


conflict_by_continent |>
  ggplot(aes(x = reorder(Continent, -total_conflict), y = total_conflict, fill = Continent)) +
  geom_col() +
  labs(
    title = "Total Conflict Incidents by Continent",
    x = "Continent",
    y = "Total Conflict Incidents",
    fill = "Continent"
  ) +
  theme_minimal() +
  scale_fill_brewer(palette = "Blues") +
  theme(legend.position = "none")
```

```{r}
#| label: fig-military-expenditure-asia 
#| layout-ncol: 2
#| fig-cap: 
#| - "Total Incident Count vs. Military Expenditures"
#| - "Total Incident Count vs. Iron and Steel Production"
#| message: false
#| warning: false


asia |>
  ggplot(aes(x = milex, y = count)) + 
  geom_point(alpha = 0.4) + 
  # geom_smooth(method = "lm") + 
  labs(
    title = "Total Incident Count vs. Military Expenditures",
    subtitle = "In Asian Countries from 1951 to 2016",
    x = "Military Expenditures (thousands in international currency)", 
    y = "Total Incident Count"
  ) + 
  theme_minimal()

asia |>
  ggplot(aes(x = irst, y = count)) + 
  geom_point(alpha = 0.4) + 
  # geom_smooth(method = "lm") + 
  labs(
    title = "Total Incident Count vs. Iron and Steel Production",
    subtitle = "In Asian Countries from 1951 to 2016",
    x = "Iron and Steel Production (thousands of tons)", 
    y = "Total Incident Count"
  ) + 
  theme_minimal()
```

@tbl-cor-table demonstrates that the correlations between incident count and military expenditures, military personnel, and iron and steel production are 0.21, 0.106, and 0.163, respectively.
While all relatively low, this finding caused us to narrow our focus onto military expenditures and iron and steel production, as those had the highest linear correlation with total incident count.

```{r}
#| label: tbl-cor-table
#| echo: false
#| tbl-cap: "Demonstrating Correlation between Total Incident Counts and Material Capabilities"

asia_cor_table <- tibble(
  `Military Capability` = c("Military Expenditures (thousands in international currency)", "Military Personnel (thousands)", "Iron and Steel Production (thousands of tons)"),
  `Correlation To Incident Count` = c(cor(asia$count, asia$milex), cor(asia$count, asia$milper), cor(asia$count, asia$irst))
)

asia_cor_table |>
  knitr::kable() |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```

<!--- classifying -->

```{r}
#| label: mutate-data 

asia <- asia |> 
  mutate(
    milex_class = case_when(
      milex < median(milex, na.rm = TRUE) ~ "low",
      .default = "high"
    ),
    
    irst_class = case_when(
      irst <= median(irst, na.rm = TRUE) ~ "low",
      .default = "high"
  )
)
```

To identify the relationship between conflict frequency and military capabilities (focusing on iron and steel production and military expenditures), we decided to do a series of visualizations using histograms.

By grouping countries into “high” and “low” military capability classes (relative to the median), the visualization explores whether military expenditures or iron and steel production are associated with a country's total incident count.

```{r}
#| label: fig-visual-milex-irst-class
#| layout-ncol: 2
#| fig-cap: 
#| - "Distribution of conflicts for countries with different military spending levels."
#| - "Distribution of conflicts for countries with different industrial power/production"

asia |>
  ggplot(aes(fill = milex_class, x = count)) + 
  geom_histogram(alpha = 0.6, binwidth = 3, position = "dodge") +
  labs(
    title = "Distribution of Incident Counts", 
    subtitle = "Grouped by higher or lower than median military expenditure", 
    y = "Number of Observations", 
    x = "Incident Counts (thousands)",
    fill = "Military Expenditure Class (relative to median)"
  ) +
  coord_cartesian(xlim = c(0, 30)) + 
  theme_minimal() +
  theme(legend.position = "bottom")

asia |>
  ggplot(aes(fill = irst_class, x = count)) + 
  geom_histogram(alpha = 0.6, binwidth = 3, position = "dodge") +
  labs(
    title = "Distribution of Incident Counts", 
    subtitle = "Grouped by higher or lower than median iron and steel production", 
    y = "Number of Observations", 
    x = "Incident Counts (thousands)",
    fill = "Iron and Steel Production Class (relative to median)"
  ) +
  coord_cartesian(xlim = c(0, 30)) + 
  theme_minimal() +
  theme(legend.position = "bottom")

```

## Methodology

As a baseline, we found the correlation coefficients between different military capabilities and the total incident count to focus our analysis on the most relevant variables (military expenditure and iron and steel production).
We generated scatterplots for said variables afterwards as we were comparing discrete numeric variables (military expenditure, iron and steel production, total incident count).
Though there were visible outliers, we could not normalize the data as several countries had 0 total incident counts, meaning we would lose a substantial number of our observations.

As a result of the extremely weak positive linear correlation we saw for all military capabilities compared to the total incident count, we decided to transform the data based off of the relatively most relevant variables (military expenditure and iron and steel production), classifying observations as either being above or below the median military expenditure as "high" and "low", respectively.
We did the same for iron and steel production.
We then generated dodged histogram visualizations for military expenditure and iron and steel production compared with total incident counts, colored by their military expenditure and iron and steel production classifications.
We chose to classify by medians rather than means due to the data's strong right skew.
Histograms fit our data as we were analyzing a discrete numerical variable, where most of the data had the same value, or around the same value.
We visualized the difference in classifications to further explore if there was a noticeable difference in total incident counts depending on countries' military expenditure and iron and steel production classifications, even if there was not a strong linear correlation.

To answer our research question, we conducted two hypothesis tests and two bootstrap simulations to determine if there was a statistically significant relationship between an Asian country's total incident number and military expenditure alongside iron and steel production respectively.
Our hypothesis tests allowed us to assess the relationship between military expenditures, iron production, and conflict frequency.
Our hypothesis tests were based on differences in medians rather than means because of how heavily skewed our data was.

The null and alternative hypotheses for our hypothesis tests are as shown:

***Military Expenditure***

**Null Hypothesis:** No statistically significant correlation exists between an Asian country’s military expenditure and total incident number.

**Alternative Hypothesis:** There is a statistically significant correlation between an Asian country’s military expenditure and total incident number.

***Iron and Steel Production***

**Null Hypothesis:** No statistically significant correlation exists between an Asian country’s iron and steel production and total incident number.

**Alternative Hypothesis:** There is a statistically significant correlation between an Asian country’s iron and steel production and total incident number.

We used bootstrap simulations to assess the strength of the relationship between an Asian country’s military expenditure, iron and steel production, and the total number of incidents.
To do this, we categorized each observation as having "high" or "low" military expenditure by comparing its value to the median military expenditure.
Given the data's strong right skew, we chose to compare medians rather than means.

In the military expenditure simulation, we calculated the median total incident count for countries with high military expenditure and low military expenditure, averaging the results across resamples to obtain multiple estimates for each group.
We followed the same procedure for the iron and steel production simulation, only evaluating iron and steel production instead of military expenditure.
Our findings from both simulations could thereby bolster our hypothesis test.

We chose against performing a linear regression model as from our data exploration, we found that there was little to no linear correlation between military capabilities and total incident counts.
A logistic regression model would not have been insightful as there were no clear binary variables in our data set and thus would have resulted in a poor fit.

## Results

#### Figures

*Figure 1*: Out of all the continents, Asia has the highest number of total conflicts at around 8100 conflicts, followed by Europe and Africa at around 3500 and 3450 respectively, and finally the Americas at around 2800 and Oceania at 500.
 

*Figure 2/Figure 3*: Both scatter plots reveal that there is an extremely weak positive linear relationship between military expenditure (milex) or iron and steel production (irst) and total incident number.
In both visualizations, most of the total incident counts remain around 0 to 10, with milex mainly clustering between 0 to 5 \* 10\^7 thousands of international currency and iron and steel production mainly between 0 to 1.25 \* 10\^5 thousands of tons.

*Figure 4/Figure 5*: The visualization reveals there is a highly right-skewed unimodal relationship between both an Asian country’s military expenditure (milex) and iron and steel production (irst) with its total incident count.
However, it appears that countries with milex or irst above the median tend to be the outliers, as slightly more countries with higher milex or irst compared to lower milex or irst respectively have greater incident counts. 

#### Hypothesis Test

```{r}
#| label: hypothesis-testing-milex
#| message: false
asia_hypothesis <- asia
set.seed(44)

null_distribution_milex <- asia_hypothesis |>
  specify(response = count, explanatory = milex_class) |>
  hypothesize(null = "independence") |>
  generate(reps = 1000, type = "permute") |>
  calculate(stat = "diff in means", order = c("high", "low"))

observed_median_milex <- asia_hypothesis |>
  specify(response = count, explanatory = milex_class) |>
  calculate(stat = "diff in medians", order = c("high", "low")) |>
  pull(stat)

null_distribution_milex |>
  get_p_value(obs_stat = observed_median_milex, direction = "two-sided") |>
  rename("p-value of milex" = "p_value") |>
  knitr::kable() |>
  kable_styling(full_width = FALSE, position = "left", bootstrap_options = c("striped", "hover", "condensed"))

null_distribution_milex |>
  get_confidence_interval()  |>
  rename(`Lower Bound for milex` = "lower_ci", `Upper Bound for milex` = "upper_ci") |>
  knitr::kable() |> 
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover", "condensed"))
```

```{r}
#| label: hypothesis-testing-irst
#| message: false

set.seed(44)

null_distribution_irst <- asia_hypothesis |>
  specify(response = count, explanatory = irst_class) |>
  hypothesize(null = "independence") |>
  generate(reps = 1000, type = "permute") |>
  calculate(stat = "diff in means", order = c("high", "low"))

observed_median_irst <- asia_hypothesis |>
  specify(response = count, explanatory = irst_class) |>
  calculate(stat = "diff in medians", order = c("high", "low")) |>
  pull(stat)

null_distribution_irst |>
  get_p_value(obs_stat = observed_median_irst, direction = "two-sided") |>
  rename("p-value of irst" = "p_value") |>
  knitr::kable() |>
  kable_styling(full_width = FALSE, position = "left", bootstrap_options = c("striped", "hover", "condensed"))


null_distribution_irst |>
  get_confidence_interval() |>
  rename(`Lower Bound for irst` = "lower_ci", `Upper Bound for irst` = "upper_ci") |>
  knitr::kable() |> 
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover", "condensed"))
```

#### Bootstrap Test for Military Expenditure and Conflict Frequency

```{r}
#| label: fig-explorea
#| fig-cap: "Bootstrap Test for Military expenditure and Conflict Frequency"
#| message: false
#| warning: false


#milex: Military expenditure.
#milper: Military personnel
#irst: Iron and Steel production

set.seed(44) 

threshold_milex <- median(asia$milex, na.rm = TRUE) 

# Bootstrap medians
bootstrap_medians_milex <- replicate(1000, {
  sampled_data_milex <- asia |>
    sample_frac(replace = TRUE) 
  # Resample with replacement
  
  medians_milex <- sampled_data_milex |>
    group_by(milex_class) |>
    summarize(median_count = median(count, na.rm = TRUE), .groups = "drop")
  
  # Return the medians for both groups as a tibble
bootstrap_medians_df1 <- tibble(
    high = medians_milex$median_count[medians_milex$milex_class == "high"],
    low = medians_milex$median_count[medians_milex$milex_class == "low"]
  )
}, simplify = FALSE) 

bootstrap_medians_df1 <- bind_rows(bootstrap_medians_milex)
```

```{r}
#| label: bootstrap-milex-tibble

bootstrap_medians_avg_milex <- tibble(
  `(avg high milex) incident count median` = mean(as.numeric(bootstrap_medians_df1$high), na.rm = TRUE),
  `(avg low milex) incident count median` = mean(as.numeric(bootstrap_medians_df1$low), na.rm = TRUE),
)

bootstrap_medians_avg_milex |>
  knitr::kable() |>
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover", "condensed"))  
```

#### Bootstrap Test for Iron Production and Conflict Frequency

```{r}
#| label: fig-exploreb
#| fig-cap: "Bootstrap Test for Iron Production and Conflict Frequency "
#| message: false
#| warning: false

set.seed(44) 

threshold_irst <- median(asia$irst, na.rm = TRUE) 

# Bootstrap medians
bootstrap_medians_irst <- replicate(1000, {
  sampled_data_irst <- asia |>
    sample_frac(replace = TRUE) 
  # Resample with replacement
  
  medians_irst <- sampled_data_irst |>
    group_by(irst_class) |>
    summarize(median_count = median(count, na.rm = TRUE), .groups = "drop")
  
  # Return the medians for both groups as a tibble
bootstrap_medians_df2 <- tibble(
    high = medians_irst$median_count[medians_irst$irst_class == "high"],
    low = medians_irst$median_count[medians_irst$irst_class == "low"]
  )
}, simplify = FALSE) 

bootstrap_medians_df2 <- bind_rows(bootstrap_medians_irst)

```

```{r}
#| label: irst-bootstrap-tibble

bootstrap_medians_avg_irst <- tibble(
  `(avg high irst) incident count median` = mean(as.numeric(bootstrap_medians_df2$high), na.rm = TRUE),
  `(avg low irst) incident count median` = mean(as.numeric(bootstrap_medians_df2$low), na.rm = TRUE),
  )

bootstrap_medians_avg_irst |>
  knitr::kable() |>
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover", "condensed"))  
```

**Hypothesis Test:** This analysis suggests that there is no statistically significant correlation between military expenditures or iron and steel production and the number of conflicts entered.
Specifically, the p-value for our military expenditures hypothesis was 0.998, with a confidence interval for the difference in median number of conflicts being  \[-0.70.
0.64\].
For our iron and steel production hypothesis, the p-value was 0.998 with a confidence interval of \[-0.65, 0.68\].
These large p-values indicate that we fail to reject the null hypothesis.

**Bootstrap Simulation:** This conclusion is supported by our bootstrap simulation, which found that the average median number of conflicts for countries with high or low military expenditures (and for those with high or low iron and steel production) was consistently 1.
The lack of difference between medians aligns with our confidence intervals, reinforcing the likelihood that any observed differences were due to random chance.

## Discussion

All aspects of our analysis suggest an essentially nonexistent linear correlation between the number of conflicts and the military capabilities we analyzed, disproving our hypothesis.
While the low correlation coefficients for material capabilities and total incident count suggest a very weak positive linear relationship between military expenditures, iron and steel production, and incident count, this relationship is minimal.
Visualizations, such as Figures 2 and 3, support this, showing that most data points cluster around 0 to 10 incident counts for both variables.
The histogram in Figure 4, which compares total incident counts with the binary categories milex_class (military expenditure classification) and irst_class (iron and steel production classification), further illustrates this pattern.
Although the proportion of highly militarized countries slightly increases with the number of conflicts, the differences are small, and these are likely outliers.
Within each bin (e.g., the number of conflicts entered), countries with "low" and "high" military expenditures are present in similar amounts, and the same trend is observed for iron production in Figure 5.
These figures also highlight the heavily right-skewed nature of our data.

When using binary variables to differentiate between “high” and “low” military expenditures and iron production, the statistical insignificance becomes more evident.
In our bootstrap simulation, the median number of conflicts remained constant (1) regardless of military expenditure or iron and steel production classification.
Additionally, extremely high p-values led us to fail to reject the null hypotheses, reinforcing the conclusion that there is no statistically significant correlation between military expenditures or iron and steel production and total incident count.

Overall, with such a heavily right-skewed data set and a median equal to the lower values, such models fail to provide a good fit for our data.

**Limitations/Improvements**

A major limitation we faced in this project was that we were constrained to the methods we had learned in class.
Although we wanted to employ a logistic or linear regression model, we found that both wouldn’t have been representative of our data – from the correlation coefficient and our visualization, we could see that there was not a significant linear correlation between the military capabilities we were analyzing and the total incident counts, and so linear regression did not make sense.
This was emphasized by Figure 2 and 3, as there wasn’t a linear relationship between any of the variables we were inspecting and the total incident count. 

Similarly, because we only had access to the total number of incident conflicts countries participated in rather than chose to not participate in, we didn’t have a clear binary categorical variable to conduct a logistic regression either.
We tried to the best of our ability to still use techniques from class by adding columns, classifying the military expenditure or iron and steel production as being above or below their medians, respectively, but our interpretations would have been improved by just using different models or techniques that were better fit to our data set. 

**Potential Issues**

Although we looked solely at military capabilities, there are other variables that cannot necessarily be quantified into numerical or categorical variables that could have influenced our data set that we did not consider.
For instance, if certain countries are in certain peace treaties or alliances that bind them to acting or not acting in conflict even if they could have.
Because it’s difficult to quantify the number of conflicts that countries refuse to enter, there also comes the possibility that a country could have a higher incident count, but in reality, is less likely to participate in conflict.
In the end, we found little correlation between expenditures, resource capabilities, and number of conflicts due to the complex nature of geopolitics. 

**Future Work**

Due to the sheer size of this data set, we had to narrow our focus onto just Asia, but in the future, we should expand our analysis to include other continents, not just Asia.
It could have also been interesting to explore how the continent with the lowest amount of incident rates compared with Asia.
Considering that the data sets contained a lot more variables than just military capabilities, a future direction could also involve looking at those other variables that we did not initially consider, such as GDP or country population. 

Additionally, the severity of conflicts is not reflected in our analysis.
Our data includes a metric detailing the severity of each conflict, but we chose not to include an “average severity of conflicts” variable when summing the number of conflicts for each country because this statistic is very similar for all countries regardless of their number of conflicts.\
