---
title: Influence of National Resources on Country Conflict
subtitle: Report
execute:
  echo: false
---

```{r}
#| label: load-pkgs
#| message: false
#| echo: false
library(tidyverse)
library(readxl)
library(countrycode) 
library(tidymodels)
library(openintro)
library(ggcorrplot)
library(broom)
library(kableExtra)

#main dataset abt resources

country_resource <- read_csv("data/NMC.csv")

#main dataset abt conflict

country_conflict <- read_csv("data/Conflict.csv")

```

This study examines the relationship between a country's national material capabilities — such as military expenditures, personnel, and iron — and its frequency engaging in conflict.
By analyzing data from the Correlates of War project, which tracks national capabilities and militarized interstate disputes, this research aims to identify whether increases in these capabilities correlate with higher frequencies of interstate conflict.
The findings seek to enhance our understanding of how shifts in state power influence international stability and conflict dynamics.

## Introduction and Data

<!--- cleaning -->

```{r}
#| label: taking-out-unnecessary-variables
#| echo: false

country_resource <- country_resource |>
  select(-ccode, -version)

country_conflict <- country_conflict |>
  select(-ccode, -version)
```

```{r}
#| label: join-datasets
#| echo: false
country_join <- country_conflict |>
  full_join(country_resource, by = join_by(
      stabb == stateabb, 
      styear == year
    )
  )

number_of_disputes_df <- country_join |>
  group_by(stabb, styear) |>
  summarize(count = n(), .groups = "drop")


```

```{r}
#| label: filter-dataset
#| echo: false
country_join_edit <- country_join |>
  left_join(number_of_disputes_df, by = c("stabb" = "stabb", "styear" = "styear"))

country_join_edit <- country_join |>
  left_join(number_of_disputes_df, by = c("stabb" = "stabb", "styear" = "styear"))

country_join_shorter <- country_join_edit |>
  select(stabb, styear, count, milex, milper, irst, cinc) |>
  distinct() |>
  filter(styear>1950)
```

```{r}
#| label: continents 
#| message: false
#| warning: false
#| echo: false

# Manually addded countried from the COW dataset, COW codes do not match universal country codes so there were many NAs using the country code package to use the continents 

# cow code to country name --> to actual ISO code --> continents 

#stabb_values <- country_join_edit[["stabb"]] isolating COW code
#unique_stabb_values <- unique(stabb_values) removing duplicates; 219 countries unique 



CountryName <- data.frame(
  COWn = c("ZIM", "ZAN", "ZAM", "YUG", "YPR", "YEM", "YAR", "WSM", "WRT", "VEN",
           "VAN", "UZB", "USA", "URU", "UKR", "UKG", "UGA", "UAE", "TUV", "TUS",
           "TUR", "TUN", "TRI", "TON", "TOG", "TKM", "THI", "TAZ", "TAW", "TAJ", 
           "SYR", "SWZ", "SWD", "SWA", "SVG", "SUR", "SUD", "STP", "SSD", "SRI", 
           "SPN", "SOM", "SOL", "SNM", "SLV", "SLU", "SLO", "SKN", "SIN", "SIE", 
           "SIC", "SEY", "SEN", "SAX", "SAU", "SAL", "SAF", "RWA", "RVN", "RUS", 
           "ROM", "ROK", "QAT", "PRK", "POR", "POL", "PNG", "PMA", "PHI", "PER", 
           "PAR", "PAP", "PAN", "PAL", "PAK", "OMA", "NTH", "NOR", "NIR", "NIG", 
           "NIC", "NEW", "NEP", "NAU", "NAM", "MZM", "MYA", "MSI", "MOR", "MON", 
           "MOD", "MNG", "MNC", "MLT", "MLI", "MLD", "MEX", "MEC", "MAW", "MAS", 
           "MAL", "MAG", "MAD", "MAC", "MAA", "LUX", "LIT", "LIE", "LIB", "LES", 
           "LEB", "LBR", "LAT", "LAO", "KZK", "KYR", "KUW", "KOS", "KOR", "KIR", 
           "KEN", "JPN", "JOR", "JAM", "ITA", "ISR", "IRQ", "IRN", "IRE", "INS", 
           "IND", "ICE", "HUN", "HSG", "HSE", "HON", "HKG", "HAN", "HAI", "GUY", 
           "GUI", "GUA", "GRN", "GRG", "GRC", "GNB", "GMY", "GHA", "GFR", "GDR",
           "GAM", "GAB", "FSM", "FRN", "FIN", "FIJ", "EU", "ETM", "ETH", "EST", 
           "ERI", "EQG", "EGY", "ECU", "DRV", "DRC", "DOM", "DMA", "DJI", "DEN", 
           "CZR", "CZE", "CYP", "CUB", "CRO", "COS", "CON", "COM", "COL", 
           "CHN", "CHL", "CHA", "CEN", "CDI", "CAP", "CAO", "CAN", "CAM", 
           "BUL", "BUI", "BRU", "BRA", "BOT", "BOS", "BOL", "BNG", "BLZ", "BLR",
           "BHU", "BHM", "BFO", "BEN", "BEL", "BAV", "BAR", "BAH", "BAD", "AZE", 
           "AUS", "AUL", "AUH", "ARM", "ARG", "ANG", "AND", 
           "ALG", "ALB", "AFG"),
  
  
  Country = c("Zimbabwe", "Zanzibar", "Zambia", "Yugoslavia/Serbia",
              "Yemen People's Republic", "Yemen", "Yemen Arab Republic", 
              "Samoa", "Wuerttemburg", "Venezuela", "Vanuatu", "Uzbekistan", 
              "United States of America", "Uruguay", "Ukraine", 
              "United Kingdom", "Uganda", "United Arab Emirates", "Tuvalu", 
              "Tuscany", "Turkey", "Tunisia", "Trinidad and Tobago", "Tonga", 
              "Togo", "Turkmenistan", "Thailand", "Tanzania", "Taiwan",
              "Tajikistan", "Syria", "Switzerland", "Sweden", "Swaziland", 
              "St. Vincent and the Grenadines", "Suriname", "Sudan",
              "Sao Tome and Principe", "South Sudan", "Sri Lanka", "Spain", 
              "Somalia", "Solomon Islands", "San Marino", "Slovenia", 
              "St. Lucia", "Slovakia", "St. Kitts and Nevis", "Singapore", 
              "Sierra Leone", "Two Sicilies", "Seychelles", "Senegal", 
              "Saxony", "Saudi Arabia", "El Salvador", "South Africa",
              "Rwanda", "Republic of Vietnam", "Russia", "Romania", 
              "South Korea", "Qatar", "North Korea", "Portugal", "Poland", 
              "Papua New Guinea", "Parma", "Philippines", "Peru", "Paraguay",
              "Papal States", "Panama", "Palau", "Pakistan", "Oman", 
              "Netherlands", "Norway", "Niger", "Nigeria", "Nicaragua", 
              "New Zealand", "Nepal", "Nauru", "Namibia", "Mozambique",
              "Myanmar", "Marshall Islands", "Morocco", "Mongolia", "Modena", 
              "Montenegro", "Monaco", "Malta", "Mali", "Moldova", 
              "Mexico", "Mecklenburg Schwerin", "Malawi", "Mauritius", 
              "Malaysia", "Madagascar", "Maldives", "Macedonia", "Mauritania", 
              "Luxembourg", "Lithuania", "Liechtenstein", "Libya", "Lesotho",
              "Lebanon", "Liberia", "Latvia", "Laos", "Kazakhstan", 
              "Kyrgyzstan", "Kuwait", "Kosovo", "Korea", "Kiribati", 
              "Kenya", "Japan", "Jordan", "Jamaica", "Italy", "Israel", "Iraq",
              "Iran", "Ireland", "Indonesia", "India", "Iceland", "Hungary",
              "Hesse Grand Ducal", "Hesse Electoral", "Honduras", "Hong Kong", 
              "Hanover", "Haiti", "Guyana", "Guinea", "Guatemala", "Grenada", 
              "Georgia", "Greece", "Guinea-Bissau", "Germany", "Ghana", 
              "German Federal Republic", "German Democratic Republic", 
              "Gambia", "Gabon", "Federated States of Micronesia", "France", 
              "Finland", "Fiji", "European Community/Union", "East Timor",
              "Ethiopia", "Estonia", "Eritrea", "Equatorial Guinea", 
              "Egypt", "Ecuador", "Vietnam", "Democratic Republic of the Congo",
              "Dominican Republic", "Dominica", "Djibouti", "Denmark", 
              "Czech Republic", "Czechoslovakia", "Cyprus", "Cuba", 
              "Croatia", "Costa Rica", "Congo", "Comoros", "Colombia", "China",
              "Chile", "Chad", "Central African Republic", "Ivory Coast", 
              "Cape Verde", "Cameroon", "Canada", "Cambodia", "Bulgaria", 
              "Burundi", "Brunei", "Brazil", "Botswana", 
              "Bosnia and Herzegovina", "Bolivia", "Bangladesh", "Belize", 
              "Belarus", "Bhutan", "Bahamas", "Burkina Faso", "Benin", 
              "Belgium", "Bavaria", "Barbados", "Bahrain", "Baden",
              "Azerbaijan", "Austria", "Australia", "Austria-Hungary", 
              "Armenia", "Argentina", "Angola", "Andorra", "Algeria", 
              "Albania", "Afghanistan")
)

country_join_countrynames <- country_join_shorter |>
  left_join(CountryName, by = c("stabb" = "COWn")) 


# country column to get the ISO codes and continents
country_join_continents <- country_join_countrynames |>
  mutate(
    # join for country code conversion
    Country_code = countrycode(Country, "country.name", "iso3c"),
    # convert the ISO country code to continent
    Continent = countrycode(Country_code, "iso3c", "continent")
  ) |>
  filter(!is.na(Country)) |>
  filter(!is.na(Continent))

#fliter by asia 
asia <- country_join_continents |>
  filter(Continent == "Asia")

```

The primary motivation for this research is to explore the relationship between national material capabilities and a country’s likelihood of being involved in conflicts.
Understanding how changes in a nation’s military capabilities — such as military expenditures and natural resources like iron — affect its involvement in international disputes is important for assessing political stability.

The data used in our analysis comes from the Correlates of War (COW) project, which consists of two datasets that apply to our research- National Material Capabilities and Militarized Interstate Disputes.
The former dataset includes annually collected data on military power indicators for each nation, such as military expenditures, military personnel, etc.
The latter records instances where states threaten, display, or use force against one another every year from 1816 to 2014.

After joining these two datasets we mutated our dataset to include the variables: “count” (summing up the total incident count for a given country in a given year), “milex_class” (classifying the military expenditure as being either above or below the median military expenditure), and “irst_class”(classifying the iron and steel production as being either above or below the median iron and steel Production).
In preparing the data, we addressed several challenges.
Variables irrelevant to our analysis were removed, and we were forced to correct the unique country codes used in the dataser as they didn’t align with standard naming conventions.
These codes were manually mapped to their respective countries (so as to not introduce a new dataset), and we classified nations adding new variables, country and continent, to narrow the scope of analysis.
\

@tbl-dict provides an overview of relevant variables used in this analysis:

```{r}
#| label: tbl-dict
#| echo: false
#| tbl-cap: "Excerpt of Data Dictionary"


variable_info <- tibble::tribble(
  ~Variable,      ~Definition,                                                                 ~Type,
  "country",      "Nation of observation",                         "categorical",
  "continent",    "Geographic area of the country (e.g., Asia)",                 "categorical",
  "count",        "Number of incidents of conflict",                    "numerical",
  "milex",        "Military expenditures in local currency(?) (in thousands)",      "numerical",
  "irst",         "Iron and Steel production levels in metric tons",             "numerical"
)

variable_info |>
kable()
```

Our research question is as follows: How do changes in national material capabilities influence the frequency of a country's conflicts, focusing on the continent of Asia?

We hypothesize that countries with higher military expenditures and personnel are more likely to participate in armed conflicts.

This research aligns with a broader theme of examining how data can explain political dynamics, provide insights into state behavior, and potentially help predict future conflicts.
In analyzing the intersection of politics and data, this study aims to tell a story about how power, in its different forms, shapes international relations.

In order to narrow our search for proper analyses we decided to focus on the continent with the highest propensity for conflict.
@fig-high-prop-conflict identifies Asia as the continent with the highest propensity for conflict.

```{r}
#| label: fig-high-prop-conflict 
#| fig-cap: "Continent with the Highest Propensity for Conflict"
#| message: false
#| warning: false
#| echo: false

conflict_by_continent <- country_join_continents |>
  group_by(Continent) |>
  summarize(total_conflict = sum(count, na.rm = TRUE)) |>
  arrange(desc(total_conflict))


conflict_by_continent |>
  ggplot(aes(x = reorder(Continent, -total_conflict), y = total_conflict, fill = Continent)) +
  geom_col() +
  labs(
    title = "Total Conflict Incidents by Continent",
    x = "Continent",
    y = "Total Conflict Incidents",
    fill = "Continent"
  ) +
  theme_minimal() +
  scale_fill_brewer(palette = "Blues") 
```

Explanatory analysis of the response variable count showed a variation in conflict frequency across countries, with some nations consistently engaging in dispute.
Patterns in military expenditure and iron and steel production were also examined in order to determine the relationships between material capabilities and conflict frequency.

We aimed to explore whether it would be worthwhile to investigate a potential relationship, so we visualized periods with heightened levels of both conflict and military capabilities, with a particular focus on iron.
Figures 2 and 3 shows an increase in activity around the same periods.

```{r}
#| label: fig-military-expenditure-asia 
#| layout-ncol: 2
#| fig-cap: 
#| - "Total Incident Count vs. Military Expenditures"
#| - "Total Incident Count vs. Iron and Steel Production"
#| message: false
#| warning: false


asia |>
  ggplot(aes(x = milex, y = count)) + 
  geom_point() + 
  # geom_smooth(method = "lm") + 
  labs(
    title = "Total Incident Count vs. Military Expenditures",
    subtitle = "In Asian Countries from 1951 to 2016",
    x = "Military Expenditures (thousands)", 
    y = "Total Incident Count"
  ) + 
  theme_minimal()

asia |>
  ggplot(aes(x = irst, y = count)) + 
  geom_point() + 
  # geom_smooth(method = "lm") + 
  labs(
    title = "Total Incident Count vs. Iron and Steel Production",
    subtitle = "In Asian Countries from 1951 to 2016",
    x = "Iron and Steel Production (thousands of tons)", 
    y = "Total Incident Count"
  ) + 
  theme_minimal()
```

We decided to do a correlation tibble to showm how key material capabilities such as military expenditure, personnel, and iron production are related to conflict frequency.

@tbl-cor-table demonstrates that the correlations between incident count and military expenditures, military personnel, and iron and steel production, are 0.21, 0.106, and 0.163 respectively.
While all relatively low, this finding caused us to narrow our focus onto military expenditures and iron and steel production, as those had the highest linear correlation with total incident count.

```{r}
#| label: tbl-cor-table
#| echo: false
#| tbl-cap: "Demonstrating Correlation between Total Incident Counts and Material Capabilities"

asia_cor_table <- tibble(
  `Military Capability` = c("Military Expenditures (thousands)", "Military Personnel (thousands)", "Iron and Steel Production (thousands of tons)"),
  `Correlation To Incident Count` = c(cor(asia$count, asia$milex), cor(asia$count, asia$milper), cor(asia$count, asia$irst))
)

asia_cor_table |>
  knitr::kable() |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```

<!--- classifying -->

```{r}
#| label: mutate-data 

asia <- asia |> 
  mutate(
    milex_class = case_when(
      milex < median(milex, na.rm = TRUE) ~ "low",
      .default = "high"
    ),
    
    irst_class = case_when(
      irst <= median(irst, na.rm = TRUE) ~ "low",
      .default = "high"
  )
)
```

To identify the relationship between conflict frequency and military capabilities we decided to do a series of visualizations using histograms.
Focusing on the Military Expenditures and Iron production as they had the highest correlation within the set.

By grouping countries into “high” and “low” military capability classes (relative to the median), the visualization explores whether military spending is associated with the frequency of incidents.

```{r}
#| label: fig-visual-milex-irst-class
#| layout-ncol: 2
#| fig-cap: 
#| - "Distribution of conflicts for countries with different military spending levels."
#| - "Distribution of conflicts for countries with different industrial power/production"

asia |>
  ggplot(aes(fill = milex_class, x = count)) + 
  geom_histogram(alpha = 0.6, binwidth = 3, position = "dodge") +
  labs(
    title = "Distribution of Incident Counts", 
    subtitle = "Grouped by higher than median military expenditure or lower than median military expenditure", 
    y = "Number of Observations", 
    x = "Incident Counts (thousands)",
    fill = "Military Expenditure Class (relative to median)"
  ) +
  coord_cartesian(xlim = c(0, 30)) + 
  theme_minimal() +
  theme(legend.position = "bottom")

asia |>
  ggplot(aes(fill = irst_class, x = count)) + 
  geom_histogram(alpha = 0.6, binwidth = 3, position = "dodge") +
  labs(
    title = "Distribution of Incident Counts", 
    subtitle = "Grouped by higher or lower than median iron and steel production", 
    y = "Number of Observations", 
    x = "Incident Counts (thousands)",
    fill = "Iron and Steel Production Class (relative to median)"
  ) +
  coord_cartesian(xlim = c(0, 30)) + 
  theme_minimal() +
  theme(legend.position = "bottom")

```

## Methodology

Due to the extensive nature of the dataset (covering numerous countries), we narrowed the focus of our project to Asia, which after visualizing the data, we found to be the continent with the highest Incident Count.

To answer our research question, we conducted two hypothesis tests and two bootstrap simulations to determine if there was a statistically significant relationship between an Asian total incident number and military expenditure alongside iron and steel production respectively.
Our hypothesis tests allowed us to assess the relationship between military expenditures, iron production and conflict frequency.

The null and alternative hypothesis for our hypothesis tests are shown below:

**Military Expenditure**

**Null Hypothesis:** No statistically significant relationship exists between an Asian country’s military expenditure and total incident number.

**Alternative Hypothesis:** There is a statistically significant relationship between an Asian country’s military expenditure and total incident number.

**Iron and Steel Production**

**Null Hypothesis:** No statistically significant relationship exists between an Asian country’s iron and steel production and total incident number.

**Alternative Hypothesis:** There is a statistically significant relationship between an Asian country’s iron and steel production and total incident number.

We used bootstrap simulations to assess the strength of the relationship between an Asian country’s military expenditure, iron and steel production, and the total number of incidents.
To do this, we categorized each observation as having "high" or "low" military expenditure by comparing its value to the median military expenditure.
Given the data's strong right skew, we chose to compare medians rather than means.

In the military expenditure simulation, we calculated the median total incident count for countries with high military expenditure and low military expenditure, averaging the results across resamples to obtain multiple estimates for each group.
We followed the same procedure for the iron and steel production simulation, only evaluating iron and steel production instead of military expenditure.
Our findings from both simulations could thereby bolster our hypothesis test.

We chose against performing a linear regression model as from our data exploration, we found that there was little to no linear correlation between military capabilities and total incident counts.
A logistic regression model would not have been insightful as there were no clear binary variables in our data set and thus would have resulted in a poor fit.

## Results

#### Bootstrap Test for Military expenditure and Conflict Frequency

```{r}
#| label: fig-explorea
#| fig-cap: "Bootstrap Test for Military expenditure and Conflict Frequency"
#| message: false
#| warning: false


#milex: Military expenditure.
#milper: Military personnel
#irst: Iron and Steel production

set.seed(44) 

threshold_milex <- median(asia$milex, na.rm = TRUE) 

# Bootstrap medians
bootstrap_medians_milex <- replicate(1000, {
  sampled_data_milex <- asia |>
    sample_frac(replace = TRUE) 
  # Resample with replacement
  
  medians_milex <- sampled_data_milex |>
    group_by(milex_class) |>
    summarize(median_count = median(count, na.rm = TRUE), .groups = "drop")
  
  # Return the medians for both groups as a tibble
bootstrap_medians_df1 <- tibble(
    high = medians_milex$median_count[medians_milex$milex_class == "high"],
    low = medians_milex$median_count[medians_milex$milex_class == "low"]
  )
}, simplify = FALSE) 

bootstrap_medians_df1 <- bind_rows(bootstrap_medians_milex)
```

```{r}
#| label: bootstrap-milex-tibble

bootstrap_medians_avg_milex <- tibble(
  `(avg high milex) incident count median` = mean(as.numeric(bootstrap_medians_df1$high), na.rm = TRUE),
  `(avg low milex) incident count median` = mean(as.numeric(bootstrap_medians_df1$low), na.rm = TRUE),
)

bootstrap_medians_avg_milex |>
  knitr::kable() |>
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover", "condensed"))  
```

#### Bootstrap Test for Iron Production and Conflict Frequency

```{r}
#| label: fig-exploreb
#| fig-cap: "Bootstrap Test for Iron Production and Conflict Frequency "
#| message: false
#| warning: false

set.seed(44) 

threshold_irst <- median(asia$irst, na.rm = TRUE) 

# Bootstrap medians
bootstrap_medians_irst <- replicate(1000, {
  sampled_data_irst <- asia |>
    sample_frac(replace = TRUE) 
  # Resample with replacement
  
  medians_irst <- sampled_data_irst |>
    group_by(irst_class) |>
    summarize(median_count = median(count, na.rm = TRUE), .groups = "drop")
  
  # Return the medians for both groups as a tibble
bootstrap_medians_df2 <- tibble(
    high = medians_irst$median_count[medians_irst$irst_class == "high"],
    low = medians_irst$median_count[medians_irst$irst_class == "low"]
  )
}, simplify = FALSE) 

bootstrap_medians_df2 <- bind_rows(bootstrap_medians_irst)

```

```{r}
#| label: irst-bootstrap-tibble

bootstrap_medians_avg_irst <- tibble(
  `(avg high irst) incident count median` = mean(as.numeric(bootstrap_medians_df2$high), na.rm = TRUE),
  `(avg low irst) incident count median` = mean(as.numeric(bootstrap_medians_df2$low), na.rm = TRUE),
  )

bootstrap_medians_avg_irst |>
  knitr::kable() |>
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover", "condensed"))  
```

#### Hypothesis Test

The shows \*\*\*, supports our hypothesis

$$
H_{0}: μ_{sample} = μ_{bootstrap}: no
$$

$$
H_{A}: μ_{sample} ≠ μ_{bootstrap}: yes
$$

```{r}
#| label: hypothesis-testing-milex
#| message: false
asia_hypothesis <- asia
set.seed(44)

null_distribution_milex <- asia_hypothesis |>
  specify(response = count, explanatory = milex_class) |>
  hypothesize(null = "independence") |>
  generate(reps = 1000, type = "permute") |>
  calculate(stat = "diff in means", order = c("high", "low"))

observed_median_milex <- asia_hypothesis |>
  specify(response = count, explanatory = milex_class) |>
  calculate(stat = "diff in medians", order = c("high", "low")) |>
  pull(stat)

null_distribution_milex |>
  get_p_value(obs_stat = observed_median_milex, direction = "two-sided") |>
  rename("p-value of milex" = "p_value") |>
  knitr::kable() |>
  kable_styling(full_width = FALSE, position = "left", bootstrap_options = c("striped", "hover", "condensed"))

null_distribution_milex |>
  get_confidence_interval()  |>
  rename("lower_ci of milex" = "lower_ci", "upper_ci of milex" = "upper_ci") |>
  knitr::kable() |> 
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover", "condensed"))

null_distribution_irst <- asia_hypothesis |>
  specify(response = count, explanatory = irst_class) |>
  hypothesize(null = "independence") |>
  generate(reps = 1000, type = "permute") |>
  calculate(stat = "diff in means", order = c("high", "low"))

observed_median_irst <- asia_hypothesis |>
  specify(response = count, explanatory = irst_class) |>
  calculate(stat = "diff in medians", order = c("high", "low")) |>
  pull(stat)

null_distribution_irst |>
  get_p_value(obs_stat = observed_median_irst, direction = "two-sided") |>
  rename("p-value of irst" = "p_value") |>
  knitr::kable() |>
  kable_styling(full_width = FALSE, position = "left", bootstrap_options = c("striped", "hover", "condensed"))


null_distribution_irst |>
  get_confidence_interval() |>
  rename("lower_ci of irst" = "lower_ci", "upper_ci of irst" = "upper_ci") |>
  knitr::kable() |> 
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover", "condensed"))
```

All aspects of our analysis point to an essentially-nonexistent correlation between the number of conflicts and any of the variables in our dataset.
When including all countries, the correlation matrix, which includes correlation coefficients such as 0.21 and 0.16 when checking for correlation with count, shows positive, linear, and very weak relationships.
Visualizations such as @fig-military-expenditure-asia and @fig-iron-production-asia demonstrate this well; the linear regressions are positive, but they fit the data very poorly.
When using a binary variable to differentiate between “high” and “low” military expenditure and iron production, a similar pattern of statistical insignificance emerges.
In general, with such a skewed dataset with the median being equal to the low, models will not fit well at all.

The histogram @fig-visual-milex-class comparing incident count and binary categories of `milex_class` and `irst_class` reflects this as well.
Although the proportion of highly militarized countries in each bin increases as the number of conflicts increases, the difference is small.
In each bin (i.e., how many conflicts entered), there are similar amounts of countries with “low” and “high” military expenditure.
The same occurs for iron production in @fig-visual-irst-class.

So far, our visualizations have given a surface-level insight regarding the influence of material capabilities and military expenditures.
To quantify these observations, we decided to employ bootstrapping and hypothesis testing.
Bootstrapping to find a difference in medians was not a useful exercise; the medians of both the “low” group and the “high” group for `milex_class` and `irst_class` were 1.
This points to a further bolstering of the claim that there is no correlation between either metric and number of conflicts entered.
Using hypothesis testing, the null hypotheses read as follows: “There is no correlation between military expenditures and number of conflicts entered” and “There is no correlation between iron production and number of conflicts entered.” For military expenditures, the confidence interval for the difference in means of number of conflicts is \[-0.70.
0.64\] with a p-value of 0.998, and for iron production, the confidence interval for the difference in means of number of conflicts is \[-0.65, 0.68\] with a p-value of 0.998.
These staggering p-values combined with confidence intervals centered around 0 confirm the null hypothesis, or the lack of correlation between the metrics and number of conflicts, with extreme confidence.
As a result, we reject the alternative hypotheses that there are correlations between military expenditures, iron production, and number of conflicts.
The p-values and confidence intervals provide a useful metric to quantify the lack of correlation, and confidently answer the research question with a resounding no – there is no relationship.

## Discussion

In this section you’ll include a summary of what you have learned about your research question along with statistical arguments supporting your conclusions.
In addition, discuss the limitations of your analysis and provide suggestions on ways the analysis could be improved.
Any potential issues pertaining to the reliability and validity of your data and appropriateness of the statistical analysis should also be discussed here.
Lastly, this section will include ideas for future work.
