---
title: Influence of National Resources on Country Conflict
subtitle: Report
execute:
  echo: true
---

AT THE END --> set echo to false to not display the code cells ; only plots, words, and messages; cite gpt in code cells  
DOI citations or manually write refernces and citations > not a requirement 

This study examines the relationship between a country's national material capabilities—such as military expenditures, personnel, and iron—and its frequency engaging in conflict.
By analyzing data from the Correlates of War project, which tracks national capabilities and militarized interstate disputes, the research aims to identify whether increases in these capabilities correlate with higher frequencies of interstate conflict.
The findings seek to enhance our understanding of how shifts in state power influence international stability and conflict dynamics.

## Introduction and data

This section includes an introduction to the project motivation, data, and research question.
Describe the data and definitions of key variables.
It should also include some exploratory data analysis.
All of the EDA won’t fit in the paper, so focus on the EDA for the response variable and a few other interesting variables and relationships.

Question:

How do changes in national material capabilities influence the frequency of a country's conflicts, focusing on the continent of Asia?

Motivation:

The primary motivation for this research is to explore the relationship between national material capabilities and a country’s likelihood of being involved in conflicts.
Understanding how changes in a nation’s military capabilities—such as military expenditures and natural resources like iron—affect its involvement in international disputes is important for assessing political stability.
This research aligns with the broader theme of examining how data can explain political dynamics, provide insights into state behavior, and potentially help predict future conflicts.

In analyzing the intersection of politics and data, this study aims to tell a story about how power, in its different forms, shapes international relations.

Source:

The data used in this analysis comes from the Correlates of War (COW) project, which is widely used by Political Scientists.
This project involves two datasets for our research:

National Material Capabilities (Dataset 1) This dataset includes data on military power indicators such as military expenditures, military personnel, energy consumption, and total population for various countries.

The data is collected annually, providing a comprehensive look at the material capabilities of each nation over time.

Militarized Interstate Disputes (Dataset 2) This dataset records instances where states threaten, display, or use force against one another.
The data includes details such as the start and end dates of conflicts, number of combatants, casualties, and the outcomes of these disputes.

This dataset spans from 1816 to 2014 and provides insight into the frequency and intensity of conflicts between countries.

## Methodology

This section includes a brief description of your analysis process.
Explain the reasoning for the types of analyses you do, exploratory, inferential, or modeling.
If you’ve chosen to do inference, make sure to include a justification for why that inferential approach is appropriate.
If you’ve chosen to do modeling, describe the model(s) you’re fitting, predictor variables considered for the model including any interactions.
Additionally, show how you arrived at the final model by describing the model selection process, interactions considered, variable transformations (if needed), assessment of conditions and diagnostics, and any other relevant considerations that were part of the model fitting process.

Our project explored the relationship between military conflicts and various factors, including military expenditure, personnel, iron and steel production, and national capabilities. Given the extensive dataset, which covered numerous countries, we initially faced challenges with clarity and focus. To streamline our analysis, we narrowed the scope to Asia, which emerged as the continent with the highest conflict frequency during our initial exploration.

To prepare the data, we addressed several challenges. Variables irrelevant to our analysis were removed, and we worked with unique country codes that didn’t align with standard naming conventions. These codes were manually mapped to their respective countries, and we classified nations by continent. This refinement allowed us to focus specifically on Asian countries. Using this dataset, we created visualizations to identify patterns in conflict frequency, iron production, and military strength metrics. For example, we examined the periods with the highest levels of conflict and iron production and determined which countries contributed most to these metrics. Early visualizations were cluttered due to the large number of countries, so we filtered the data to focus on the top 15 countries in terms of conflict frequency and iron production to improve clarity.

We also investigated correlations between national capabilities—such as military expenditure and conflict frequency to identify potential predictors. To assess the impact of these variables, we used regression modeling, with predictor variables.

For our analysis, we utilized several key R libraries to streamline data manipulation, modeling, and visualization. The tidyverse suite was useful in providing tools for filtering and summarizing data and ggplot for crafting visualizations. We used readxl to import Excel datasets into R. The countrycode package helped map country codes from the dataset to their corresponding country names, helping us to group data by continent and eventually determine the focus on Asian countries. For modeling and diagnostics, we relied on tidymodels for regression analysis.

Combinging data preparation, visualization, statistical analysis, and modeling, we found meaningful trends and relationships that highlight how national capabilities influence military conflicts different countries in Asia. This comprehensive approach allowed us to derive insights into the multifaceted drivers of geopolitical events.


## Results

This is where you will discuss your overall finding and describe the key results from your analysis.
The goal is not to interpret every single element of an output shown, but instead to address the research questions, using the interpretations to support your conclusions.
Focus on the variables that help you answer the research question and that provide relevant context for the reader.

## Discussion

In this section you’ll include a summary of what you have learned about your research question along with statistical arguments supporting your conclusions.
In addition, discuss the limitations of your analysis and provide suggestions on ways the analysis could be improved.
Any potential issues pertaining to the reliability and validity of your data and appropriateness of the statistical analysis should also be discussed here.
Lastly, this section will include ideas for future work.

# Data Exploration

```{r}
#| label: load-pkgs
#| message: false
library(tidyverse)
library(readxl)
library(countrycode) #this isn't working for me? //you have to download it
library(tidymodels)
library(openintro)

#main dataset abt resources

country_resource <- read_csv("data/NMC.csv")

#main dataset abt conflict

country_conflict <- read_csv("data/Conflict.csv")

```

# Cleaning Data

```{r}
#| label: join-datasets

country_resource <- country_resource |>
  select(-ccode, -version)

country_conflict <- country_conflict |>
  select(-ccode, -version)
```

```{r}
country_join <- country_conflict |>
  full_join(country_resource, by = join_by(
      stabb == stateabb, 
      styear == year
    )
  )

number_of_disputes_df <- country_join |>
  group_by(stabb, styear) |>
  summarize(count = n(), .groups = "drop")


```

```{r}
country_join_edit <- country_join |>
  left_join(number_of_disputes_df, by = c("stabb" = "stabb", "styear" = "styear"))
#ggplot(country_join_edit, aes(x = milex, y = count)) +
  #geom_point() 
```

# Adding Continents

```{r}
#| label: continents 
#| message: false
#| warning: false

#adding countries, the COW codes do not match the universal country codes so there are many nas using the countrycode package
# cow code to country name --> to actual ISO code --> continents 
#stabb_values <- country_join_edit[["stabb"]] isolating COW code
#unique_stabb_values <- unique(stabb_values) removing duplicates 219 countries noted 


country_join_edit <- country_join |>
  left_join(number_of_disputes_df, by = c("stabb" = "stabb", "styear" = "styear"))

country_join_shorter <- country_join_edit |>
  select(stabb, styear, count, milex, milper, irst, cinc) |>
  distinct() |>
  filter(styear>1950)

CountryName <- data.frame(
  COWn = c("ZIM", "ZAN", "ZAM", "YUG", "YPR", "YEM", "YAR", "WSM", "WRT", "VEN", "VAN", 
           "UZB", "USA", "URU", "UKR", "UKG", "UGA", "UAE", "TUV", "TUS", "TUR", "TUN", "TRI", 
           "TON", "TOG", "TKM", "THI", "TAZ", "TAW", "TAJ", "SYR", "SWZ", "SWD", "SWA", "SVG", 
           "SUR", "SUD", "STP", "SSD", "SRI", "SPN", "SOM", "SOL", "SNM", "SLV", "SLU", "SLO", 
           "SKN", "SIN", "SIE", "SIC", "SEY", "SEN", "SAX", "SAU", "SAL", "SAF", "RWA", "RVN", 
           "RUS", "ROM", "ROK", "QAT", "PRK", "POR", "POL", "PNG", "PMA", "PHI", "PER", "PAR", 
           "PAP", "PAN", "PAL", "PAK", "OMA", "NTH", "NOR", "NIR", "NIG", "NIC", "NEW", "NEP", 
           "NAU", "NAM", "MZM", "MYA", "MSI", "MOR", "MON", "MOD", "MNG", "MNC", "MLT", "MLI", 
           "MLD", "MEX", "MEC", "MAW", "MAS", "MAL", "MAG", "MAD", "MAC", "MAA", "LUX", "LIT", 
           "LIE", "LIB", "LES", "LEB", "LBR", "LAT", "LAO", "KZK", "KYR", "KUW", "KOS", "KOR", 
           "KIR", "KEN", "JPN", "JOR", "JAM", "ITA", "ISR", "IRQ", "IRN", "IRE", "INS", "IND", 
           "ICE", "HUN", "HSG", "HSE", "HON", "HKG", "HAN", "HAI", "GUY", "GUI", "GUA", "GRN", 
           "GRG", "GRC", "GNB", "GMY", "GHA", "GFR", "GDR", "GAM", "GAB", "FSM", "FRN", "FIN", 
           "FIJ", "EU", "ETM", "ETH", "EST", "ERI", "EQG", "EGY", "ECU", "DRV", "DRC", "DOM", 
           "DMA", "DJI", "DEN", "CZR", "CZE", "CYP", "CUB", "CRO", "COS", "CON", "COM", "COL", 
           "CHN", "CHL", "CHA", "CEN", "CDI", "CAP", "CAO", "CAN", "CAM", "BUL", "BUI", "BRU", 
           "BRA", "BOT", "BOS", "BOL", "BNG", "BLZ", "BLR", "BHU", "BHM", "BFO", "BEN", "BEL", 
           "BAV", "BAR", "BAH", "BAD", "AZE", "AUS", "AUL", "AUH", "ARM", "ARG", "ANG", "AND", 
           "ALG", "ALB", "AFG"),
  Country = c("Zimbabwe", "Zanzibar", "Zambia", "Yugoslavia/Serbia", "Yemen People's Republic", 
              "Yemen", "Yemen Arab Republic", "Samoa", "Wuerttemburg", "Venezuela", "Vanuatu", 
              "Uzbekistan", "United States of America", "Uruguay", "Ukraine", "United Kingdom", 
              "Uganda", "United Arab Emirates", "Tuvalu", "Tuscany", "Turkey", "Tunisia", 
              "Trinidad and Tobago", "Tonga", "Togo", "Turkmenistan", "Thailand", "Tanzania", 
              "Taiwan", "Tajikistan", "Syria", "Switzerland", "Sweden", "Swaziland", 
              "St. Vincent and the Grenadines", "Suriname", "Sudan", "Sao Tome and Principe", 
              "South Sudan", "Sri Lanka", "Spain", "Somalia", "Solomon Islands", "San Marino", 
              "Slovenia", "St. Lucia", "Slovakia", "St. Kitts and Nevis", "Singapore", 
              "Sierra Leone", "Two Sicilies", "Seychelles", "Senegal", "Saxony", "Saudi Arabia", 
              "El Salvador", "South Africa", "Rwanda", "Republic of Vietnam", "Russia", "Romania", 
              "South Korea", "Qatar", "North Korea", "Portugal", "Poland", "Papua New Guinea", 
              "Parma", "Philippines", "Peru", "Paraguay", "Papal States", "Panama", "Palau", 
              "Pakistan", "Oman", "Netherlands", "Norway", "Niger", "Nigeria", "Nicaragua", 
              "New Zealand", "Nepal", "Nauru", "Namibia", "Mozambique", "Myanmar", "Marshall Islands", 
              "Morocco", "Mongolia", "Modena", "Montenegro", "Monaco", "Malta", "Mali", "Moldova", 
              "Mexico", "Mecklenburg Schwerin", "Malawi", "Mauritius", "Malaysia", "Madagascar", 
              "Maldives", "Macedonia", "Mauritania", "Luxembourg", "Lithuania", "Liechtenstein", 
              "Libya", "Lesotho", "Lebanon", "Liberia", "Latvia", "Laos", "Kazakhstan", 
              "Kyrgyzstan", "Kuwait", "Kosovo", "Korea", "Kiribati", "Kenya", "Japan", "Jordan", 
              "Jamaica", "Italy", "Israel", "Iraq", "Iran", "Ireland", "Indonesia", "India", "Iceland", 
              "Hungary", "Hesse Grand Ducal", "Hesse Electoral", "Honduras", "Hong Kong", 
              "Hanover", "Haiti", "Guyana", "Guinea", "Guatemala", "Grenada", "Georgia", "Greece", 
              "Guinea-Bissau", "Germany", "Ghana", "German Federal Republic", "German Democratic Republic", 
              "Gambia", "Gabon", "Federated States of Micronesia", "France", "Finland", "Fiji", 
              "European Community/Union", "East Timor", "Ethiopia", "Estonia", "Eritrea", "Equatorial Guinea", 
              "Egypt", "Ecuador", "Vietnam", "Democratic Republic of the Congo", "Dominican Republic", 
              "Dominica", "Djibouti", "Denmark", "Czech Republic", "Czechoslovakia", "Cyprus", "Cuba", 
              "Croatia", "Costa Rica", "Congo", "Comoros", "Colombia", "China", "Chile", "Chad", 
              "Central African Republic", "Ivory Coast", "Cape Verde", "Cameroon", "Canada", "Cambodia", 
              "Bulgaria", "Burundi", "Brunei", "Brazil", "Botswana", "Bosnia and Herzegovina", "Bolivia", 
              "Bangladesh", "Belize", "Belarus", "Bhutan", "Bahamas", "Burkina Faso", "Benin", "Belgium", 
              "Bavaria", "Barbados", "Bahrain", "Baden", "Azerbaijan", "Austria", "Australia", "Austria-Hungary", 
              "Armenia", "Argentina", "Angola", "Andorra", "Algeria", "Albania", "Afghanistan")
)

country_join_countrynames <- country_join_shorter |>
  left_join(CountryName, by = c("stabb" = "COWn")) 


# country column to get the ISO codes and continents
country_join_continents <- country_join_countrynames |>
  mutate(
    # join for country code conversion
    Country_code = countrycode(Country, "country.name", "iso3c"),
    # convert the ISO country code to continent
    Continent = countrycode(Country_code, "iso3c", "continent")
  ) |>
  filter(!is.na(Country))
country_join_continents


```

# Iron
This scatterplot shows the relationship between iron and steel production (in thousands of tons) and the incident count across countries, categorized by continent. The data suggests that most countries with low iron and steel production have a wide range of incident counts, while countries with high production tend to have fewer incidents, concentrated at lower counts. Notably, Asia and Europe dominate the high-production range, while other continents exhibit more dispersed patterns in lower production ranges.
```{r}
#| label: graph-Iron-and-IncidentCount
#| message: false
#| warning: false
ggplot(country_join_continents, aes(x = irst, y = count, color = Continent)) +
  geom_point() + 
   labs(title = "Iron and steel production (thousands of tons) vs Incident Count by Country",
       x = "Iron and steel production (thousands of tons)",
       y = "Count") +
  theme_minimal()

```

# Which Continent has the Highest Propensity for Conflict?

```{r}
#| label: high-prop-conflict 
#| message: false
#| warning: false

conflict_by_continent <- country_join_continents |>
  group_by(Continent) |>
  summarize(total_conflict = sum(count, na.rm = TRUE)) |>
  arrange(desc(total_conflict))


conflict_by_continent |>
  ggplot(aes(x = reorder(Continent, -total_conflict), y = total_conflict, fill = Continent)) +
  geom_col() +
  labs(
    title = "Total Conflict Incidents by Continent",
    x = "Continent",
    y = "Total Conflict Incidents",
    fill = "Continent"
  ) +
  theme_minimal() +
  scale_fill_brewer(palette = "Blues") #make it pretty 
```

Focus on the continent with the highest propensity for conflict, which is Asia.
```{r}
#| label: period-conflicts
#| message: false
#| warning: false
#| layout-ncol: 2

#fliter by asia 
asia <- country_join_continents |>
  filter(Continent == "Asia")


asia |>
  group_by(styear) |>
  summarize(total_conflict = sum(count, na.rm = TRUE)) |>
  ggplot(aes(x = styear, y = total_conflict)) +
  geom_line() +
  labs(
    title = "Trends in Conflict Incidents in Asia",
    x = "Year",
    y = "Total Conflict Incidents"
  ) +
  theme_minimal()


#fliter by asia 
asia <- country_join_continents |>
  filter(Continent == "Asia")


# Confict and Iron production start around the same years.

asia |>
  group_by(styear) |>
  summarize(total_iron_production = sum(irst, na.rm = TRUE)) |>
  ggplot(aes(x = styear, y = total_iron_production)) +
  geom_line() +
  labs(
    title = "Trends in Iron and steel production (thousands of tons) in Asia",
    x = "Year",
    y = "Total Conflict Incidents around the same years as conflict"
  ) +
  theme_minimal()


```



Countries in Asia that lead in Incident Count and Iron Production

```{r}
#| label: top-countries-conflict-iron
#| message: false
#| warning: false
#| layout-ncol: 2

top_15_countries_in_asia <- asia |>
  group_by(stabb) |>
  summarize(total_conflict = sum(count, na.rm = TRUE)) |>
  arrange(desc(total_conflict)) |>
  head(15) |>
  pull(stabb)
asia_15 <- asia |>
  filter(stabb %in% top_15_countries_in_asia)


asia_15 |>
  group_by(stabb) |>
  summarize(total_conflict = sum(count, na.rm = TRUE)) |>
  arrange(desc(total_conflict))


asia_15 |>
  group_by(stabb) |>
  summarize(total_conflict = sum(count, na.rm = TRUE)) |>
  ggplot(aes(x = reorder(stabb, -total_conflict), y = total_conflict)) +
  geom_col(fill = "steelblue") +
  labs(
    title = "Conflict Incidents by Country in Asia",
    x = "Country",
    y = "Total Conflict Incidents"
  ) +
  theme_minimal()
asia_15 |>
  group_by(stabb) |>
  summarize(total_iron_production = sum(irst, na.rm = TRUE)) |>
  arrange(desc(total_iron_production))

asia_15 |>
  group_by(stabb) |>
  summarize(total_iron_production = sum(irst, na.rm = TRUE)) |>
  ggplot(aes(x = reorder(stabb, -total_iron_production), y = total_iron_production)) +
  geom_col(fill = "steelblue") +
  labs(
    title = "Iron Production by Country in Asia",
    x = "Country",
    y = "Total Iron Production"
  ) +
  theme_minimal()
```

# Explore

Relations between material capabilities and conflict frequency.

```{r}
#| label: top-countries-conflict-iron-cor
#| message: false
#| warning: false
#milex: Military expenditure.
#milper: Military personnel
#irst: Iron and Steel production
#cinc: Composite Index of National Capability.
#count: Frequency of conflicts.
asia |>
  select(milex, milper, irst, cinc, count) |>
  cor(use = "complete.obs")
# calc correlations; no missing val accounted 
```

1: Perfect positive correlation (as one variable increases, the other always increases).
0: No correlation.
-1: Perfect negative correlation (as one variable increases, the other always decreases).

```{r}
#| label: countries-conflict-iron-model
#| message: false
#| warning: false

country_conflict_incident_model <- linear_reg() |>
  fit(count ~ milex + milper + irst + cinc, data = asia)

tidy(country_conflict_incident_model)

```


```{r}
#| label: disp-mean
country_join_continents |>
  summarize(mean = mean(count))

```


# Logistic Regression Model for more than 50 disputes in Asia

```{r}
#| label: log-disp-asia 
asia_log_raw <- country_join_continents |>
  mutate(
    disp_above_50 = if_else(count >= 50, "yes", "no"),
    disp_above_50 = fct_relevel(disp_above_50, "yes", "no")
  ) |>
  filter(Continent == "Asia")

asia_fit <- logistic_reg() |>
  fit(disp_above_50 ~ irst + milex + milper, data = asia_log_raw)

tidy(asia_fit)
```

$$
\widehat{disp~above~50} = 5.077 - (1.353 * 10^{-6}) * irst - (2.528 * 10^{-9}) * milex - (6.643 * 10^{-6}) * milper
$$

