---
title: States Resources compared to Likelihood of being involved in conflict 
subtitle: Proposal
---

```{r}
#| label: load-pkgs
#| message: false
library(tidyverse)
```

# Data 1

## Introduction and data

-   Identify the source of the data.

    Correlates of War

    -   Dataset 1: National Material Capabilities

    -   Dataset 2: Militarized Interstate Disputes

-   State when and how it was originally collected (by the original data curator, not necessarily how you found the data).

    The Correlates of War project generally facilitates the collaboration and collecting of data about the history of wars and conflicts among different states.
    It was started in 1963 at the University of Michigan by J.
    David Singer.
    This data set was prepared by scraping information from sources from each country, such as the US Department of Commerce and Bureau of the Census or The Japan Year Book Office.

-   Write a brief description of the observations.

    Observations dataset 1: Each observation includes a breakdown of annual data of a certain country in a particular year.
    This observation includes information about military expenditures, the number of military personnel, energy consumption, total population, etc.

    Observations dataset 2: This dataset contains information on wars between countries.
    Each observation is for a specific year and country, which breaks down the number of conflicts in which a country threatens/displays/uses force against one or more states.
    The observation includes the start and end dates that the observation is tracking between, number of combatants, number of casualties and outcomes, etc.
    The dataset includes observations between 1816 and 2014.
    Other information included in the observations are the start and end date of incident, duration, etc.

## Research question

-   A well formulated research question.
    (You may include more than one research question if you want to receive feedback on different ideas for your project. However, one per data set is required.)

    How do changes in national material capabilities influence the likelihood of a country being involved in conflict?

-   A description of the research topic along with a concise statement of your hypotheses on this topic.

    This research topic explores the varying levels of impact each national capability indicator has on a country's likelihood of entering conflict.
    It also reflects which national capability indicator has a greater correlation with the likelihood of conflict involvement.

    Thus, our hypothesis is:

    -   Countries with higher military expenditures and personnel are more likely to participate in armed conflicts.

-   Identify the types of variables in your research question.
    Categorical?
    Quantitative?

    -   military expenditures - quantitative

        military personnel - quantitative

        likelihood of entering conflict - quantitative

## Glimpse of data

```{r}
#| label: load-data-1

#main dataset abt resources

country_resource <- read_csv("data/NMC.csv")
glimpse(country_resource)

#main dataset abt conflict

country_conflict <- read_csv("data/Conflict.csv")
glimpse(country_conflict)


```

# Data 2

## Introduction and data

-   Identify the source of the data.

-   State when and how it was originally collected (by the original data curator, not necessarily how you found the data).

-   Write a brief description of the observations.

## Research question

-   A well formulated research question. (You may include more than one research question if you want to receive feedback on different ideas for your project. However, one per data set is required.)
-   A description of the research topic along with a concise statement of your hypotheses on this topic.
-   Identify the types of variables in your research question. Categorical? Quantitative?

## Glimpse of data

```{r}
#| label: load-data-2


```
