---
title: JADA Proposals
author: Allison Yang, Daniel Matten, Jacob Ominsky, Ashley Deleon
---

```{r}
#| label: load-pkgs
#| message: false
library(tidyverse)
library(readxl)
```

# Data 1

## Introduction and data

This data comes from the Equity in Athletics Data Analysis, which is created by the U.S. Department of Education. Institutions of higher education that participate in federal student aid programs are obligated to submit athletics data by the Equity in Athletics Disclosure Act, and that data appears in the Equity in Athletics Disclosure database as well as this data. The data is collected every year as schools fill out the mandatory survey, and it includes information such as spending on recruiting, head coaching salary, and financial aid for each school.


## Research question

-   A well formulated research question. (You may include more than one research question if you want to receive feedback on different ideas for your project. However, one per data set is required.)

What is the correlation between athletic expenditures for football programs and their performance metrics?

Are there regional differences in football expenditures among various colleges and universities? How do these differences relate to team performance?

How do investments in football facilities (stadiums, training facilities) correlate with overall athletic department budgets and expenditures?

What is the relationship between coaching salaries in football and overall athletic expenditures? How do these salaries compare to those in other sports?

-   A description of the research topic along with a concise statement of your hypotheses on this topic.
Our research topic focuses on analyzing the EADA's annual, which includes spending on recruiting, coaching salaries, and financial aid. We predict that there is a positive relationship between spending and football success in NCAA Division 1 thanks to the impact of additional scholarships and nicer facilities. However, we predict diminishing marginal returns on that spending -- the difference in spending between the top teams is likely insignificant.

-   A statement about why this question is important.

Spending in college football is more rampant than ever with the injection of NIL money into the sport, so schools without such a longstanding football tradition (like SEC schools) might rethink their spending habits.

-   Identify the types of variables in your research question. Categorical? Quantitative?
Besides the variables identifying the sport name and institution name (and associated variables such as state), all of the variables are numerical -- spending, number of participants, etc. Some rows are labeled as "logical" because the row is filled with NAs. Some schools do not have specific sports or specific teams, so expenditure data cannot be provided for them.


## Glimpse of data

```{r}
#| label: load-data-1
data <- read_xlsx("Sport_Data_2022 FOOTBALL.xlsx")
glimpse(data)
```

# Data 2

## Introduction and data

-   Identify the source of the data.

    Correlates of War

    -   Dataset 1: National Material Capabilities

    -   Dataset 2: Militarized Interstate Disputes

-   State when and how it was originally collected (by the original data curator, not necessarily how you found the data).

    The Correlates of War project generally facilitates the collaboration and collecting of data about the history of wars and conflicts among different states.
    It was started in 1963 at the University of Michigan by J.
    David Singer.
    This data set was prepared by scraping information from sources from each country, such as the US Department of Commerce and Bureau of the Census or The Japan Year Book Office.

-   Write a brief description of the observations.

    Observations dataset 1: Each observation includes a breakdown of annual data of a certain country in a particular year.
    This observation includes information about military expenditures, the number of military personnel, energy consumption, total population, etc.

    Observations dataset 2: Each observation includes a breakdown of nations and the conflict in which one or more states threaten, display, or use force against one or more other states between 1816 and 2014.
    Each observation includes the start and end date of incident, duration, etc.

## Research question

-   A well formulated research question.
    (You may include more than one research question if you want to receive feedback on different ideas for your project. However, one per data set is required.)

    How do changes in national material capabilities influence the likelihood of a country being involved in conflict?

-   A description of the research topic along with a concise statement of your hypotheses on this topic.

    This research topic explores the varying levels of impact each national capability indicator has on a country's likelihood of entering conflict.
    It also reflects which national capability indicator has a greater correlation with the likelihood of conflict involvement.

    Thus, our hypothesis is:

    -   Countries with higher military expenditures and personnel are more likely to participate in armed conflicts.

-   Identify the types of variables in your research question.
    Categorical?
    Quantitative?

    -   military expenditures - quantitative

        military personnel - quantitative

        likelihood of entering conflict - quantitative

## Glimpse of data

```{r}
#| label: load-data-2

#main dataset abt resources

country_resource <- read_csv("NMC.csv")
glimpse(country_resource)

#main dataset abt conflict

country_conflict <- read_csv("Conflict.csv")
glimpse(country_conflict)


```